Fluent statistical computing interfaces for biological data analysis

### 2020-10-27 response/response.Rmd

I have expanded this section to reflect these comments.

### Chapter 2 comments

Most of the comments pertain to the use and learning of the **plyranges** package. I believe they are best addressed through the **plyranges** documentation and vignette. I have amended the thesis to include the package vignette as an appendix.

### Chapter 5 comments

>  Pg 77: “non-liner” should be “non-linear”

>  Pg 78: “produces” should be “produce”

> Pg 78: Could you add a reference for PCA?

Amended.
> Pg 80 (also in 81): The way “tSNE” was written in “this essentially turns t-SNE” looks very different
than previously written stylistically. Could this be converted to the way you were originally referring to
it?

> Pg 87: “inn” should be “in”

> Pg 88: Inside Figure 5.1A and B, could you have liminal label which plot is which i.e. t-SNE vs tour
view. Also, could axis labels be added? Maybe also add this information to the Figure legend? – Pg 89 – same comment

I think this is not obvious in the static figure but is obvious in the video and when using the **liminal** package that the embedding view is always on the left, and the tour view is on the right (it will be the view that is animated). I think the suggestion of titling the views is a good one and will place that into the software in future. Regarding axis labels, I don't think they make sense for the tour since it's really a linear combination of multiple variables (there's an option to switch on a biplot view for the tour). I am also reluctant to include them for t-SNE because they can be misleading for the reasons mentioned in the chapter, however I do think this would be good to have as an option for **liminal** in future.

> Pg 90: I believe there should be a period after “tree structure data (figure 5.3)” instead of a comma

> Pg 94: Can you comment on the scRNA-seq example where you only pick first 5 PCs that explained only 20% of the variance. Or more generally, what is the trade-off on the liminal view by picking more or less PCs along with a NLDR method. Do you have recommendations for the user on how many
PCs to pick for and optimal liminal view?

I have amended section 5.5.3 to provide more details about why the first five components were selected, and how to choose variables to tour.

> Pg 94: Can you comment on the limits of the liminal view in terms of number of observations that
one can visualize in a given tour? Have you thought about strategies for making this scale with more observations?
– Ah, I see you touched on this at the end of the discussion of the capture. I think this is a really important point that would be greatly appreciated going forward.
– And it was talked about in Chapter 6 in Future Work. Incorporating this more in the middle of the chapter with tables specifying what is or is not reasonable to plot in terms of the number of observations, would be helpful.

I have amended section 5.5.3 to discuss a strategy in the specific case of cluster analysis. I have also amended the discussion of the chapter with specific work that can mitigate the need to show all of the data. In the case of tasks performed for clustering I don't think it is useful or necessary to show all points in the scatter plot when we are most interested in density and shapes. There is interesting work to be done here in future for appropriate sampling strategies or aggregations that could be done prior to producing a visualisation.

### 2020-10-27 Rmd/05-embeddings.Rmd
more details for chapter 5

To check and verify the clustering we construct a **liminal** view. 
We tour the first five PCs (approximately 20% of the variance in expression),
alongside the t-SNE view which was computed from all 25 PCs. 
Due to latency of the interface we do a weighted sample of  the rows based on
cluster membership, leaving us with approximately 10 per cent of the original
data size - 4,590 cells. Although this is not ideal, it still allows us to
get a sense of the shape of the clusters as seen from the linked video in figure \@ref(fig:mouse-01).

@@

There are some limitations
in using the **liminal** interface for larger datasets. First,
t-SNE avoids the crowding problem, points are separated into distinct regions
on the display canvas, while for the tour, points are concentrated in the
centre of the projection and become difficult to see. 
Second, as $n$ increases the animation becomes slower as there is more data
passing from the server to the client. Both of these issues could be mitigated
through alternate tour paths and displays. The recent work
by @Laa2020-wr is a promising area for further investigation, as well as
work from topological statistics [@Rieck2017-kk; @Genovese2017-iq].

>>>>

To check and verify the clustering we construct a **liminal** view. 
We tour the first five PCs (approximately 20% of the variance in expression),
alongside the t-SNE view which was computed from all 25 PCs. We have selected only the first five PCs because there is a large drop in the percentage of variance explained after the fifth component, with each component after contributing less than one percent of variance.
Consequently, increasing the number of PCs to tour would increase the dimensionality and volume of the subspace we are touring but without adding any additional signal to the view.

Due to latency of the **liminal** interface we do a weighted sample of  the rows based on
cluster membership, leaving us with approximately 10 per cent of the original
data size - 4,590 cells. Although this is not ideal, it still allows us to
get a sense of the shape of the clusters as seen from the linked video in figure \@ref(fig:mouse-01). If one was interested in performing more in-depth cluster analysis we recommend an iterative approach of removing large clusters and then re-running the **liminal** view as a way finding more granular cluster structure.

@@

There are some limitations
in using the **liminal** interface for larger datasets. First,
t-SNE avoids the crowding problem, points are separated into distinct regions
on the display canvas, while for the tour, points are concentrated in the
centre of the projection and become difficult to see. We have recently proposed a simple non-linear transformation for the tour called a sage tour that aims to fix this problem [@Laa2020-vk]. 
Second, as $n$ increases both the embedding view and tour view become harder to read due to over-plotting, while the interactivity and animation become slower as there is more data passing from the server to the client.
For the tasks we have looked at in this paper, where shape and density are important to the analyst, we think that better displays and sub-sampling strategies are more useful than being able to look at every single point on the canvas. We showed in our single cell clustering case study that doing a weighted sample based on cluster membership still allowed us to get a sense of relative cluster orientation, however there are alternative sampling approaches that could be applied, like selecting points close to the cluster centres. Alternative displays via statistical transformations could also mitigate the need to show all of the data. Recent work
by @Laa2020-wr is a promising area for further investigation, as well as
work from topological statistics [@Rieck2017-kk; @Genovese2017-iq].
